---
title: "Homework_04"
author: "Coni"
date: "October 4, 2018"
output:
  github_document:
    toc: yes
  
---

# Homework 04: Tidy data and joins

In this assignment we use `tidyr` and `dplyr`

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(gapminder)
library(tidyverse)
```


##Data Reshaping Prompts (and relationship to aggregation)

_Problem:_ You have data in one “shape” but you wish it were in another. Usually this is because the alternative shape is superior for presenting a table, making a figure, or doing aggregation and statistical analysis.

_Solution:_ Reshape your data. For simple reshaping, `gather()` and `spread()` from `tidyr` will suffice. Do the thing that is possible / easier now that your data has a new shape.

Before getting into the assignment activity, let's remember what `gather()` and `spread()` do

1. `gather()` 

It's used for making untidy data tidy. It takes multiple columns, and gathers them into key-value pairs: it makes “wide” data longer.

*Format:* `gather`(`data`, `key = "key_name"`, `value = "value"`, `...`) where:
  
  - `"key_name"` is what we want to call the column headers that became a variable, 
  
  - `"value"` is what we want to call the stacked columns that became a variable
  
  - and the last argument `...` correspond to the items we want to include or not include in the new table

Let's first have a look at the iris data frame

```{r iris}

head(iris)

```

And then take a subset of this df by slicing the first appearance of each species

```{r mini_iris}
mini_iris <- 
  iris %>%
  group_by(Species) %>% #group by species
  slice(1) #takes only the first row of each group

knitr::kable(mini_iris) #print table
```

Now we can make this untidy data tidy with gather()

```{r gather() example}
tidy_mini_iris <- mini_iris %>% 
  gather(key = flower_att, #creates a new column named flower_att that contains the measurement type
         value = measurement, #contains the value of the measurement type for each species
         -Species) #include the species name in the new table

knitr::kable(tidy_mini_iris)
```

2. `spread()` 

Takes two columns (key & value) and spreads in to multiple columns.
Useful to make “long” data wider.

```{r spread() example}
untidy_mini_iris <- spread(tidy_mini_iris,flower_att, measurement)

knitr::kable(untidy_mini_iris)
```

_Prompt:_ Activity #2

Make a tibble with one row per year and columns for life expectancy for two or more countries.

- Use knitr::kable() to make this table look pretty in your rendered homework.
- Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
head(gapminder)

(mini_gapminder <- gapminder %>%
  select(country, year, lifeExp) %>% #keep only variables we want
  filter(country == "Chile" | country == "Canada" )) # keep only countries we want

mini_gapminder %>% 
  spread(country, lifeExp) %>% #make tibble where countries are columns and years are rows
  ggplot(aes(Canada, Chile, label = year)) + #scatterplot lifeExp of Canada vs Chile
  geom_point() +
  geom_text(aes(label = year), hjust = 0, vjust = 0) #add labels and adjust position

```

##Join Prompts (join, merge, look up)

_Problem:_ You have two data sources and you need info from both in one new data object.

_Solution:_ Perform a join, which borrows terminology from the database world, specifically SQL.

_Prompt:_ Activity #1

Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins. 

Examples of a second data frame you could build:
- One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city.
- One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.

Stuck trying to get data to use as a second data frame...

```{r}
library(readxl)

#asd <- read_excel("https://github.com/Gapminder-Indicators/tfr/raw/master/tfr-by-gapminder.xlsx")
```

