---
title: "Homework_04"
author: "Coni"
date: "October 4, 2018"
output: github_document
---

# Homework 04: Tidy data and joins

In this assignment we use `tidyr` and `dplyr`

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
```


##Data Reshaping Prompts (and relationship to aggregation)

_Problem:_ You have data in one “shape” but you wish it were in another. Usually this is because the alternative shape is superior for presenting a table, making a figure, or doing aggregation and statistical analysis.

_Solution:_ Reshape your data. For simple reshaping, `gather()` and `spread()` from `tidyr` will suffice. Do the thing that is possible / easier now that your data has a new shape.

Before getting into the assignment activity, let's remember what `gather()` and `spread()` do

1. `gather()` is used for making untidy data tidy

*Format:* `gather`(`data`, `key = "key_name"`, `value = "value"`, `...`) where:
  
  - `"key_name"` is what we want to call the column headers that became a variable, 
  
  - `"value"` is what we want to call the stacked columns that became a variable
  
  - and the last argument `...` correspond to the items we want to include or not include in the new table

Let's first have a look at the iris data frame

```{r iris}

colnames(iris)

```

And then take a subset of this table by slicing the first appearance of each species

```{r mini_iris}
(mini_iris <-
  iris %>%
  group_by(Species) %>%
  slice(1))
```

Now we can make this untidy data tidy with gather()

```{r gather() example}
mini_iris %>% 
  gather(key = flower_att, #creates a new column named flower_att that contains the measurement type
         value = measurement, #contains the value of the measurement type for each species
         -Species) #we include the species name in the new table
```

