---
title: "Homework_04"
author: "Coni"
date: "October 4, 2018"
output:
  github_document:
    toc: yes
  
---

# Homework 04: Tidy data and joins

In this assignment we use `tidyr` and `dplyr`

```{r load libraries, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(gapminder)
library(tidyverse)
```


##Data Reshaping Prompts (and relationship to aggregation)

_Problem:_ You have data in one “shape” but you wish it were in another. Usually this is because the alternative shape is superior for presenting a table, making a figure, or doing aggregation and statistical analysis.

_Solution:_ Reshape your data. For simple reshaping, `gather()` and `spread()` from `tidyr` will suffice. Do the thing that is possible / easier now that your data has a new shape.

Before getting into the assignment activity, let's remember what `gather()` and `spread()` do

1. `gather()` 

It's used for making untidy data tidy. It takes multiple columns, and gathers them into key-value pairs: it makes “wide” data longer.

*Format:* `gather`(`data`, `key = "key_name"`, `value = "value"`, `...`) where:
  
  - `"key_name"` is what we want to call the column headers that became a variable, 
  
  - `"value"` is what we want to call the stacked columns that became a variable
  
  - and the last argument `...` correspond to the items we want to include or not include in the new table

Let's first have a look at the iris data frame

```{r iris}

head(iris)

```

And then take a subset of this data frame by slicing the first appearance of each species

```{r mini_iris}
mini_iris <- 
  iris %>%
  group_by(Species) %>% #group by species
  slice(1) #takes only the first row of each group

knitr::kable(mini_iris) #print table
```

Now we can make this untidy data tidy with gather()

```{r gather() example}
tidy_mini_iris <- mini_iris %>% 
  gather(key = flower_att, #creates a new column named flower_att that contains the measurement type
         value = measurement, #contains the value of the measurement type for each species
         -Species) #include the species name in the new table

knitr::kable(tidy_mini_iris)
```

2. `spread()` 

Takes two columns (key & value) and spreads in to multiple columns.
Useful to make “long” data wider.

```{r spread() example}
untidy_mini_iris <- spread(tidy_mini_iris,flower_att, measurement)

knitr::kable(untidy_mini_iris)
```

_Prompt:_ Activity #2

Make a tibble with one row per year and columns for life expectancy for two or more countries.

- Use knitr::kable() to make this table look pretty in your rendered homework.
- Take advantage of this new data shape to scatter plot life expectancy for one country against that of another.

```{r scatter plot lifeExp}
head(gapminder)

(mini_gapminder <- gapminder %>%
  select(country, year, lifeExp) %>% #keep only variables we want
  filter(country == "Chile" | country == "Canada" )) # keep only countries we want

mini_gapminder %>% 
  spread(country, lifeExp) %>% #make tibble where countries are columns and years are rows
  ggplot(aes(Canada, Chile, label = year)) + #scatterplot lifeExp of Canada vs Chile
  geom_point() +
  geom_text(aes(label = year), hjust = 0, vjust = 0) #add labels and adjust position

```

##Join Prompts (join, merge, look up)

_Problem:_ You have two data sources and you need info from both in one new data object.

_Solution:_ Perform a join, which borrows terminology from the database world, specifically SQL.

_Prompt:_ Activity #1

Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins. 

Examples of a second data frame you could build:
- One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city.
- One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.

We'll add a row with the country capitals. for this we first download the new data set and check the countries listed there match the ones in gapminder.

```{r country_capitals data set}
#get data frame with country capitals
country_capitals <- read_csv("https://github.com/icyrockcom/country-capitals/raw/master/data/country-list.csv") 

colnames(country_capitals) #inspect country_capitals column names
nrow(country_capitals) #get number of rows

colnames(gapminder) #inspect gapminder column names
nrow(gapminder$country) #number of rows

#see if the country list matches for both data sets
setdiff(gapminder$country, country_capitals$country) 
#see countries that appear in gapminder but not in country_capitals

intersect(country_capitals$country, gapminder$country) #see countries that appear in both datasets

```

We see that apparently there are some countries in `gapminder` that aren't in our new data set `country_capitals`. However we can still work with the 129 countries that are in both.

Let's join the two data sets together with `full_join()`

```{r full_join country_capitals and gapmider}
(full_join(gapminder, country_capitals, by = "country"))
```
By using `full_join()` we end up with a dataset that is gapminder plus 2 columns where we have the capital name and the type (apparently countryCapital for all)

Note that we were forced to transform factors to character vectors (because one of the datasets had the countries as factors)

Lets try `left_join()` now:

```{r left_join}
left_join(gapminder, country_capitals, by = "country")
```

And `right_join()`:

```{r right_join}
right_join(gapminder, country_capitals, by = "country")
```

This keeps the list of countries in the country_capitals data set as the new country row. Note that the number of countries is different, so one needs to be careful with this.

Let's see what happens with `inner_join()`

```{r inner_join}
inner_join(gapminder, country_capitals, by = "country")
```

This is probably the best way to do it, becasue we are keeping only the countries that we have in both datasets (we noticed before that some countries had their names spelled different in the two data sets; this way we dont have that artificial repetition at least)


