---
title: "Homework_04"
author: "Coni"
date: "October 4, 2018"
output: github_document
---

# Homework 04: Tidy data and joins

In this assignment we use `tidyr` and `dplyr`

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
```


##Data Reshaping Prompts (and relationship to aggregation)

_Problem:_ You have data in one “shape” but you wish it were in another. Usually this is because the alternative shape is superior for presenting a table, making a figure, or doing aggregation and statistical analysis.

_Solution:_ Reshape your data. For simple reshaping, `gather()` and `spread()` from `tidyr` will suffice. Do the thing that is possible / easier now that your data has a new shape.

Before getting into the assignment activity, let's remember what `gather()` and `spread()` do

- `gather()` is used for making untidy data tidy

  *Format:* gather(`data`, `key = "key_name"`, `value = "value"`, `...`) where:
  
  *"key_name"* is what we want to call the column headers that became a variable, 
  
  *"value"* is what we want to call the stacked columns that became a variable
  
  and the last argument *...* correspond to the items we want to include or not include in the new table

```{r}

```

